---
title: "Landers Module 3 Code"
author: "Demetrius Green"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_notebook: default
---

```{r include=F}
#ensure working dir is correct
#load appropriate libraries

library(tidyverse)
library(psych)
```

Read in raw data and create df!
```{r}
raw_df = read.csv("../data/landers_wk3data.csv", header = T)
glimpse(raw_df) 
```

Recast *timeStart* & *timeEnd* into POSIX format!
```{r}
raw_df = raw_df %>%
    mutate_at(vars('timeStart', 'timeEnd'), as.POSIXct)
#glimpse(raw_df) #notice the vars changed from 'factor' to 'POSIX'
```

Research assistants participated during the month of June - remove them and create
a new df. Also, *q6*was a check-response item keyed for 1.
```{r}
#research assistants participated during the month of June
#remove them and create a new df; q6 was check response item keyed for 1
clean_df = raw_df %>%
    filter_at(vars('timeStart', 'timeEnd', 'q6'), 
              any_vars(. >= "2017-07-01 00:00:00" & q6 == 1))
```

# M3.Analysis

Save the total time spent on study in seconds as a new variable. Also, recast 
*timeSpent* as a number.
```{r}
#save total time spent on study in seconds as a new variable
#recast timeSpent as num
clean_df = clean_df %>%
    mutate('timeSpent_secs' = as.numeric(difftime(
            timeEnd, timeStart, units = 'secs')))
```
\newpage

Create a histogram of timeSpent!
```{r}
#ggplot2/tidyverse syntax
ggplot(clean_df, aes(x = timeSpent_secs)) + 
    geom_histogram(stat = "bin", binwidth = 90,
                   color = "white", 
                   fill = "dark green") +
    labs(title = "Research Study Time Expenditure", 
         x = "# of seconds") +
    theme_dark()

#base R syntax
hist(clean_df$timeSpent_secs, col = "orange", 
     main = "Research Study Time Expenditure", xlab = "# of seconds")
```

Iterate table function over the 5th and 14th columns of df
```{r}
freq_tbls_list = lapply(clean_df[, c(5, 14)], table)
freq_tbls_list

#tidyverse syntax
# clean_df %>% 
#     select('q1', 'q10') %>%
#     lapply(table)
```

```{r echo = F, warning = F, message=F}
library(gridExtra)

bp1 = barplot(freq_tbls_list$q1)
bp2 = barplot(freq_tbls_list$q10)

bp1
bp2
```

Participants should have answered *q1* with a greater or equal value to what they
answered *q2* with, and they should never have answered *q2* with the same value
as what they answered *q3* with. Display a count of the number of times in *clean_df* 
that this actually happened.

```{r}
#tidyverse syntax
clean_df %>% 
  filter_at(vars('q1', 'q2', 'q3'), any_vars(q1 >= q2 & q2 != q3)) %>%
  count()
```




