---
title: 'Module 7: Data Visualization'
author: "Demetrius K. Green"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_notebook: default
---

This week, you’ll be visualizing datasets from a past project. All plots should 
be created using **ggplot2** unless otherwise specified. When linking portions of 
a ggplot(), use separate lines for each new "+" phrase. Download week3.csv from 
Project 3 into your data folder.

*Warning.* If you find that your graphs aren’t what you expect from commands you 
think are correct, you probably messed up something in your data wrangling steps.

## Data Import & Cleaning

```{r, include=FALSE}
library(tidyverse)
library(psych)
library(kableExtra)
```

Import the *week3.csv* and do the following data wrangling steps:

- convert the dataset to a tibble
- convert any data types that were not imported correctly into their correct
types using tidy functions where possible 
- relabel factors according to Table 1 (refer to worksheet)
- remove all the cases that failed the manipulation check (i.e., responded with 
anything other than a 1 to q6)
- drop q6 from the final tbl

```{r, echo = F}
week7_tbl = as_tibble(read.csv('../data/landers_wk3data.csv')) %>% 
    mutate_at(vars('timeStart', 'timeEnd'), as.POSIXct) %>% 
    mutate(
        condition = recode(condition, A = 'Block A', B = 'Block B', C = 'Control'), 
        gender = recode(gender, M = 'Male', F = 'Female')) %>%
    filter(q6 == 1) %>% 
    select(-q6)
```

\newpage

## Visualization

Create a single figure that displays a scatterplot and correlation matrix of 
q1-q5 and q7-q10 as well as density plots (i.e., the line graph version of a 
histogram). **Do not use ggplot()**.

```{r, message=F, warning=F, echo = F}
library("GGally")

week7_tbl %>% 
    select(q1:q10) %>% 
    ggpairs()
```

```{r, echo=F, include=F}
dens = apply(week7_tbl[, 5:13], 2, density)

plot(NA, xlim = range(sapply(dens, "[", "x")), ylim = range(sapply(dens, "[", "y"))) 
test = mapply(lines, dens, col = 1:length(dens))

legend("topright", legend = names(dens), fill = 1:length(dens), cex = .65)
```

\newpage

Check for differences in Q1 over time by creating a plot of "*timeStart*" on the 
x-axis and q1 on the y-axis that looks like Fig 1. 

```{r, echo = F}
week7_tbl %>% 
    ggplot(aes(timeStart, q1)) +
    geom_point() +
    labs(
        title = 'Fig 1', 
        x = 'Date of Experiments', 
        y = 'Q1 Score') +
    theme(
        plot.title = element_text(face = 'bold'))
```

\newpage 

Check for gender differences in the Q1/Q2 relationship by creating a figure that 
looks like Fig 2. Note that this is a jitter plot, so your dot positions will
probably not look the same. 

```{r, echo = F}
week7_tbl %>% 
    ggplot(aes(q1, q2)) + 
    geom_jitter(aes(color = gender))
```

\newpage

Check for gender differences in the Q1/Q2 relationship again, but this time by
creating two side-by-side plots as shown in Fig 3 (also a jitter plot)

```{r, echo = F}
week7_tbl %>%
    ggplot(aes(q1, q2)) +
    geom_jitter() + 
    facet_wrap(~gender) +
    labs(
        title = 'Fig 3', 
        x = 'Score on Q1', 
        y = 'Score on Q2') + 
    theme(
        plot.title = element_text(face = 'bold')
    )
```

\newpage

Using a single pipe, create a plot displaying the experimental time elapsed in 
seconds between "*timeStart*" and "*timeEnd*" split by gender, which looks like 
Fig 4.

```{r, echo = F}
week7_tbl %>% 
    mutate(
        'timeSpent_secs' = as.numeric(
            difftime(timeEnd, timeStart, units = 'secs')
            )
        ) %>% 
    ggplot(aes(gender, timeSpent_secs)) +
    geom_boxplot(aes(x = gender)) +
    labs(
        title = 'Fig 4', 
        x = 'Gender', 
        y = 'Time Elapsed (secs)') +
    theme(plot.title = element_text(face = 'bold'))
```

\newpage

Recreate Fig 5. Jitter may be slightly different but should be similar. The 
background of the legend is 12.5% gray. *Hint:* Tackle this figure one plus at a 
time, checking your work as you go.

```{r, echo = F}
week7_tbl %>% 
    ggplot(aes(q5, q7)) +
    geom_jitter(aes(color = condition), width = .05) +
    geom_line(aes(color = condition), 
              stat = 'summary_bin', binwidth = 3, fun.y = 'mean') +
    labs(
        fill = "Experimental Condition",
        title = 'Fig 5', 
        x = 'Score on Q5', 
        y = 'Score on Q7'
    ) +
    theme(
        plot.title = element_text(face = "bold"), 
        legend.background = element_rect(fill = '#DFDFDF'), 
        legend.position = "bottom"
    )
    
```

